Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-10-06 paper/main.tex
fiber

\note{something about how structure of data field serves as basis of equivariance in viz literature}
Besides expressing continuities, fiber bundles also provide us a way to richly express the structure of the fields in the data. Specifically, we do this by adopting Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}. Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}. In the paper, the $\dfiber$ is the cartesian cross product of the fibers for each column.

@@

\section{Artist} 
The \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the Artist object in Matplotlib\cite{hunterMatplotlib2DGraphics2007} is responsible for converting data to a prerender graphical object} is a functor from data $\mathcal{\dtotal}$ to graphic $\mathcal{\gtotal}$.}

@@

We formally define data and graphic as the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$. We develop these categories as analogues to fiber bundle $(\dtotal, \dbase, \pi, \dfiber)$, introduced in \autoref{sec:related-work:fiber-bundles}, and fiber bundle $(\gtotal, \gbase, \pi, \gfiber)$. Throughout this section, we assume that $\mathcal{E}$ is defined in terms of a fixed continuity and fiber space, for example a class of images or tables with the same schema. Similarly we assume that $\mathcal{H}$ is a stand in for a specific type of graphic, such as a scatter or line plot.

As introduced in \autoref{sec:related-work:fiber-bundles}, sets of open sets form the basis of topological spaces. We propose that we can express how the elements in our dataset are glued together by modeling that continuity as a topological space \dbase, and we express the constraints on \dbase by defining the analogues category $\mathcal{\dbase}$.

@@

The objects $U_{i}$ are a way to manage subsets of the data. For example, let there be a data value for each point along the interval $\dbase=\right[0,1\left]$. Since this function is continuous, there is a subset of the data defined along the open set $U_{a} = \right(0.2, 0.4\left)$. The open set $U_{a}$ provides a unique way of identifying that subset.

\note{maybe a small image here or when we introduce open sets showing what we mean?}

\begin{definition} 
  \label{def:category:F}
  The category $\mathcal{\dfiber}$ is a subcategory of an arbitrary category $\mathcal{C}$. $\mathcal{\dfiber}$ is a monoidal category, meaning it has one object \texttt{\dfiber} and a functor\footnote{The functor $\otimes$ is a bifunctor because it takes as input the product of two categories \cite{nlab:bifunctor}} $\otimes:\mathcal{\dfiber}\times{\mathcal{\dfiber}}\rightarrow \mathcal{\dfiber}$\cite{barrCategoryTheoryComputing}.
\end{definition}

The object $\texttt{\dfiber}$ can be an object of any arbitrary category $\mathcal{C}$, which allows $\texttt{\dfiber}$ to encode most data fields. For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.

@@

\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$. The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over open set $\U_{i}$.

  The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$. \cite{SheafMathematics2021,spanier1989algebraic}.
\end{definition}
Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, database tables that have the same schema and discrete index, as illustrated in Spivak \cite{spivakSIMPLICIALDATABASES}.

@@

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the space of continuous\footnote{or piecewise continuous}sections $\sheaf: U_{i} \rightarrow \Gamma(U_{i}, \dtotal) $ which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021}. 
    \item\textit{morphisms} bundle maps E, E, fiber to itself and is the fiber map allowable, functions F to itself forms a monoid (one object category), F and it's allowable maps form a category, monoid could be monotonic functions, group actions, extending that to bundle E, interested in E to itself, which when restricted to fiber are category maps of fiber to itself, maps from sheaf to itself, that when restricted to fiber gives you fiber to itself, given bundle map E->E, then you can show that you have a sheaf map... Given E->E, and F->F, then there's a map from one sheaf to another. bundle map induces on level of sections induces on sheafs -> part of the definition of the presheaf
    \item \textit{morphisms} E->E such that restricted to fiber it's the allowed maps, and restricted to base it's the inclusion map, sheaf to itself so is an endofunctor (category to itself)
  \end{itemize}
\end{definition}

@@

\end{definition}

% For this sake we don't do polymorphism, but we have a notion of equivalance classes of artists s.t... every parameter you want to fix is a fiber in E 
% given a nu, 
% maximal V has all the visual variables 
% nu_red()->red

>>>>

@@

\section{Artist} 
In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations. We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}. The artist \vartist converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$.}

@@

We define the categories $\mathcal{\dtotal}, \mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.  We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

As introduced in \autoref{sec:related-work:fiber-bundles}, sets of open sets form the basis of topological spaces. We propose that we can express how the elements in our data are connected to each other by modeling the data continuity as a topological space \dbase. We express subsets of the continuity, which can be mapped into subsets of the data, as open sets $U_{i}$ in \dbase. We introduce the category $\mathcal{\dbase}$ to encapsulate the subsets of the continuity and the inclusion maps that glue them together.

@@

Besides expressing continuities, fiber bundles also provide us a way to richly express the structure of the fields in the data. Specifically, we do this by adopting Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}. Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}. In the paper, the $\dfiber$ is the cartesian cross product of the fibers for each column.

The category $\mathcal{\dfiber}$ is a subcategory of an arbitrary category $\mathcal{C}$. $\mathcal{\dfiber}$ is, meaning it has one object \texttt{\dfiber} and a functor\footnote{The functor $\otimes$ is a bifunctor because it takes as input the product of two categories \cite{nlab:bifunctor}} $\otimes:\mathcal{\dfiber}\times{\mathcal{\dfiber}}\rightarrow \mathcal{\dfiber}$\cite{barrCategoryTheoryComputing}.

\begin{definition} 
  \label{def:category:F} 
  The category $\mathcal{\dfiber}$ consists of one object $\dfiber$ with maps back to itself $\monoid: \dfiber \rightarrow \dfiber$. 
\end{definition}

%F is the one object category with maps back to itself M: F\rightarrow F, fiber and all its allowed maps. morphsims in categroy C that go back to same object fixed F, we note that this is also a category w/ one object, a one object categry is a monoid

The object $\texttt{\dfiber}$ is an object of some category $\mathcal{C}$, which allows $\texttt{\dfiber}$ to encode most data fields. For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}. We note that a one object category is a monoid \cite{nlab:mo}

@@

%insteaf of trying to define gamma, write this up as a paragaraph... we use the convention \Gamma(U,E) to denote, the map from U_i to gamma U_e...defines a presheaf which is also a sheaf.... continue w/ global sections 
\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$. The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over open set $\U_{i}$. The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$. \cite{SheafMathematics2021,spanier1989algebraic}.
\end{definition}

Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, database tables that have the same schema and discrete index, as illustrated in Spivak \cite{spivakSIMPLICIALDATABASES}.

@@

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the space of continuous\footnote{or piecewise continuous}sections $\sheaf: U_{i} \rightarrow \Gamma(U_{i}, \dtotal) $ which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021}. 
    \item\textit{morphisms} allowed maps from sheaf to sheaf, when restricted to a fiber, that should be a map from a fiber to fiber.  
    \item\textit{morphisms} bundle maps E, E, fiber to itself and is the fiber map allowable, functions F to itself forms a monoid (one object category), F and it's allowable maps form a category, monoid could be monotonic functions, group actions, extending that to bundle E, interested in E to itself, which when restricted to fiber are category maps of fiber to itself, maps from sheaf to itself, that when restricted to fiber gives you fiber to itself, given bundle map E->E, then you can show that you have a sheaf map... Given E->E, and F->F, then there's a map from one sheaf to another. bundle map induces on level of sections induces on sheafs -> part of the definition of the presheaf
    \item \textit{morphisms} E->E such that restricted to fiber it's the allowed maps, and restricted to base it's the inclusion map, sheaf to itself so is an endofunctor (category to itself)
  \end{itemize}
\end{definition}

@@

\end{definition}
% For this sake we don't do polymorphism, but we have a notion of equivalance classes of artists s.t... every parameter you want to fix is a fiber in E 
% given a nu, 
% maximal V has all the visual variables 
% nu_red()->red
